Design and Implementation of Probabilistic Programming Language Anglican

### 2016-07-07 paper/paper.tex
abstract and summary

\begin{abstract}
    Anglican is a probabilistic programming system designed to interoperate
    with Clojure and other JVM languages. We describe the implementation of
    Anglican and illustrate how its design facilitates both explorative and
    industrial use of probabilistic programming.

    In these notes, I explain the rationale behind the
    macro-compiled implementation of Anglican, outline
    design choices, and describe important implementation
    aspects.
\end{abstract}

@@

>>>>

\begin{abstract}
    Anglican is a probabilistic programming system designed to interoperate
    with Clojure and other JVM languages. We introduce the
	programming language Anglican, outline our design choices,
	and discuss in depth the implementation of the Anglican language
	and runtime, including macro-based compilation, extended
	CPS-based evaluation model, and functional representations for
	probabilistic paradigms, such as distribution,
	random processes, and inference algorithms.

	We show that a probabilistic functional language can be
	implemented efficiently and integrated tightly with a
	conventional functional language with only moderate
	computational overhead. We also demonstrate how advanced
	probabilistic modeling concepts are mapped naturally
	to a functional representation.
\end{abstract}

@@

\section{Summary}

In this paper, we presented design and implementation internals
of the Probabilistic Programming System Anglican. Implementing
a language is an interesting endeavour, in particular when the
language implements a new paradigm, in this case probabilistic
programming. Functional programming is a natural complement of
probabilistic programming --- the latter allows both concise and
expressive specification of probabilistic generative models and
efficient implementation of inference algorithm.

Implementing a probabilistic language on top and in tight
integration with a functional language, Clojure, both helped
us to accomplish an ambitious goal in a short time span, and
provided important insights on structure and semantics of
probabilistic concepts incorporated in Anglican. Efficiency
and expressive power of Anglican owe to adherence to the
functional approach as much as to rich inference opportunities
by the Anglican environment.
